---
title: "Homework Assignment 3"
author: "Isabella Chittumuri"
date: "4/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Documents/Hunter College/Spring 2021/Stat 707/HW")
library(tidyverse)
library(ggplot2)
```

# 8.21 
In a regression analysis of on-the-job head injuries of warehouse laborers caused by falling objects, $Y$ is a measure of severity of the injury, $X_1$ is an index reflecting both the weight of the object and the distance it fell, and $X_2$ and $X_3$ are indicator variables for nature of head protection worn at the time of the accident, coded as follows:

|Type of Protection| $X_2$| $X_3$|
|----------|------|------|
|Hard hat |1| 0| 
|Bump cap |0| 1|
|None |0| 0|

The response function to be used in the study is $E[Y] = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3$

## a. 
Develop the response function for each type of protection category.

Hard hat ($X_2=1, X_3=0$):

$$
\begin{aligned}
E[Y] &= \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 \\
&= \beta_0 + \beta_1X_1 + \beta_2(1) + \beta_3(0) \\
&= (\beta_0 + \beta_2) + \beta_1X_1
\end{aligned}
$$

Bump hat ($X_2=0, X_3=1$):

$$
\begin{aligned}
E[Y] &= \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 \\
&= \beta_0 + \beta_1X_1 + \beta_2(0) + \beta_3(1) \\
&= (\beta_0 + \beta_3) + \beta_1X_1
\end{aligned}
$$

None ($X_2=0, X_3=0$):

$$
\begin{aligned}
E[Y] &= \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 \\
&= \beta_0 + \beta_1X_1 + \beta_2(0) + \beta_3(0) \\
&= \beta_0 + \beta_1X_1
\end{aligned}
$$

## b.
For each of the following questions, specify the alternatives $H_0$ and $H_a$ for the appropriate test: 

(1) With $X_1$ fixed, does wearing a bump cap reduce the expected severity of injury as compared with wearing no protection? 

$$
\begin{aligned}
& H_0: \beta_3 = 0 \\
& H_a: \beta_3 \ne 0
\end{aligned}
$$
In words: 

$H_0$, not wearing bump cap reduces expected severity of injury

$H_a$, wearing bump cap reduces expected severity of injury

(2) With $X_1$ fixed, is the expected severity of injury the same when wearing a hard hat as when wearing a bump cap?

$$
\begin{aligned}
& H_0: \beta_2 = \beta_3\\
& H_a: \beta_2 \ne \beta_3
\end{aligned}
$$

In words: 

$H_0$, wearing hard hat and bump cap has same expected severity of injury

$H_a$, wearing hard hat and bump cap does not have same expected severity of injury

# 8.31

## a. 
Derive the expressions for $b'_0, b'_1, \space and \space  b'_{11}$ in (8.12)

$$
\begin{aligned}
\hat{Y} &= b_0 + b_1X + b_{11}X^2\\
&= b_0 + b_1(X - \bar{X}) + b_{11}(X- \bar{X})^2 \\
&= b_0 + b_1X - b_1\bar{X} + b_{11}X^2 - 2b_{11}X\bar{X} + b_{11}\bar{X}^2 \\
&= (b_0 - b_1\bar{X} + b_{11}\bar{X}^2) + (b_1 - 2b_{11}\bar{X})X + b_{11}X^2 \\
&& Hence: \space &b'_0 = b_0 - b_1\bar{X} + b_{11}\bar{X}^2 \\
&& \space &b'_1 = b_1 - 2b_{11}\hat{X} \\
&& \space &b'_{11} = b_{11}
\end{aligned}
$$

## b.
Using (5.46) obtain the variance-covariance matrix for the regression coefficients pertaining to the original X variable in terms of the variance-covariance matrix for the regression coefficients pertaining to the transformed x variable.

(5.46) $\sigma^2[W] = \sigma^2[AY] = A\sigma^2[Y]A'$

coefficients of:
$b'_0 = [1, -\bar{X}, \bar{X}^2]$
$b'_1 = [0, 1, -2\bar{X}]$
$b'_{11} = [0, 0, 1]$

$$
A = 
\begin{bmatrix}
1 & -\bar{X} & \bar{X}^2 \\
0 & 1 & -2\bar{X} \\
0 & 0 & 1
\end{bmatrix}
$$

$$
\sigma^2[b] = 
\begin{bmatrix}
\sigma^2{b_0} & \sigma{b_0b_1} & \sigma{b_0b_2} \\
\sigma{b_1b_0} & \sigma^2{b_1} & \sigma{b_1b_2} \\
\sigma{b_2b_0} & \sigma{b_2b_1} & \sigma^2{b_2}
\end{bmatrix}
$$

$$
A' = 
\begin{bmatrix}
1 & 0 & 0 \\
-\bar{X} & 1 & 0 \\
\bar{X}^2 & -2\bar{X} & 1
\end{bmatrix}
$$

$$
A\sigma^2[b]A' = 
\begin{pmatrix}
1 & -\bar{X} & \bar{X}^2 \\
0 & 1 & -2\bar{X} \\
0 & 0 & 1
\end{pmatrix}
\begin{pmatrix}
\sigma^2{b_0} & \sigma{b_0b_1} & \sigma{b_0b_2} \\
\sigma{b_1b_0} & \sigma^2{b_1} & \sigma{b_1b_2} \\
\sigma{b_2b_0} & \sigma{b_2b_1} & \sigma^2{b_2}
\end{pmatrix}
\begin{pmatrix}
1 & 0 & 0 \\
-\bar{X} & 1 & 0 \\
\bar{X}^2 & -2\bar{X} & 1
\end{pmatrix}
$$

Matrix multiplication is associative: $ABC = (AB)C = A(BC)$

$$
\begin{aligned}
A\sigma^2[b] =
\begin{bmatrix}
\sigma^2{b_0} -\bar{X}\sigma{b_1b_0} + \bar{X}^2\sigma{b_2b_0} & \sigma{b_0b_1} -\bar{X}\sigma^2{b_1} + \bar{X}^2\sigma{b_2b_1} & \sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2}\\
\sigma{b_1b_0} - 2\bar{X}\sigma{b_2b_0}  & \sigma^2{b_1} - 2\bar{X}\sigma{b_2b_1} &  \sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2} \\
\sigma{b_2b_0} & \sigma{b_2b_1} & \sigma^2{b_2}
\end{bmatrix}
\end{aligned}
$$

$$
A\sigma^2[b]A' = 
\begin{bmatrix}
\sigma^2{b_0} -\bar{X}\sigma{b_1b_0} + \bar{X}^2\sigma{b_2b_0} & \sigma{b_0b_1} -\bar{X}\sigma^2{b_1} + \bar{X}^2\sigma{b_2b_1} & \sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2}\\
\sigma{b_1b_0} - 2\bar{X}\sigma{b_2b_0}  & \sigma^2{b_1} - 2\bar{X}\sigma{b_2b_1} &  \sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2} \\
\sigma{b_2b_0} & \sigma{b_2b_1} & \sigma^2{b_2}
\end{bmatrix}
\begin{bmatrix}
1 & 0 & 0 \\
-\bar{X} & 1 & 0 \\
\bar{X}^2 & -2\bar{X} & 1\\
\end{bmatrix}
$$

Whole Matrix:

$$
A\sigma^2[b]A' = 
\begin{bmatrix}
\sigma^2{b_0} -\bar{X}\sigma{b_1b_0} + \bar{X}^2\sigma{b_2b_0} - \bar{X}(\sigma{b_0b_1} -\bar{X}\sigma^2{b_1} + \bar{X}^2\sigma{b_2b_1}) + \bar{X}^2(\sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2}) & \sigma{b_0b_1} -\bar{X}\sigma^2{b_1} + \bar{X}^2\sigma{b_2b_1} - 2\bar{X}(\sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2}) & \sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2} \\
\sigma{b_1b_0} - 2\bar{X}\sigma{b_2b_0} -\bar{X}(\sigma^2{b_1} - 2\bar{X}\sigma{b_2b_1}) + \bar{X}^2(\sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2}) &  \sigma^2{b_1} - 2\bar{X}\sigma{b_2b_1} - 2\bar{X}(\sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2}) & \sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2} \\
\sigma{b_2b_0} - \bar{X}(\sigma{b_2b_1}) + \bar{X}^2(\sigma^2{b_2}) & \sigma{b_2b_1}-2\bar{X}(\sigma^2{b_2}) & \sigma^2{b_2}
\end{bmatrix}
$$
Separated First Column:

$$
(A\sigma^2[b]A')_{1} = 
\begin{pmatrix}
\sigma^2{b_0} -\bar{X}\sigma{b_1b_0} + \bar{X}^2\sigma{b_2b_0} - \bar{X}(\sigma{b_0b_1} -\bar{X}\sigma^2{b_1} + \bar{X}^2\sigma{b_2b_1}) + \bar{X}^2(\sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2}) \\
\sigma{b_1b_0} - 2\bar{X}\sigma{b_2b_0} -\bar{X}(\sigma^2{b_1} - 2\bar{X}\sigma{b_2b_1}) + \bar{X}^2(\sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2}) \\
\sigma{b_2b_0} - \bar{X}(\sigma{b_2b_1}) + \bar{X}^2(\sigma^2{b_2})
\end{pmatrix}
$$
Separated Second Column:

$$
(A\sigma^2[b]A')_{2} = 
\begin{bmatrix}
\sigma{b_0b_1} -\bar{X}\sigma^2{b_1} + \bar{X}^2\sigma{b_2b_1} - 2\bar{X}(\sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2}) \\
\sigma^2{b_1} - 2\bar{X}\sigma{b_2b_1} - 2\bar{X}(\sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2}) \\
\sigma{b_2b_1}-2\bar{X}(\sigma^2{b_2}) 
\end{bmatrix}
$$

Separated Third Column:

$$
(A\sigma^2[b]A')_{3} = 
\begin{bmatrix}
\sigma{b_0b_2} -\bar{X}\sigma{b_1b_2} +  \bar{X}^2\sigma^2{b_2} \\
\sigma{b_1b_2} - 2\bar{X}\sigma^2{b_2} \\
\sigma^2{b_2}
\end{bmatrix}
$$
After simplifications:

$$
\begin{aligned}
& \sigma(b'_0) = \sigma^2_0 - 2\bar{X}\sigma_{01} + 2\bar{X}^2\sigma_{02} + \bar{X}^2\sigma^2_1 - 2\bar{X}^3\sigma_{12} + \bar{X}^4\sigma^2_2 \\
& \sigma(b'_1) = \sigma^2_1 - 4\bar{X}\sigma_{12} + 4\bar{X}^2\sigma^2_2 \\
& \sigma(b'_2) = \sigma^2_2 \\
& \sigma(b'_0,b'_1) = \sigma_{01} - 2\bar{X}\sigma_{02} + 3\bar{X}^2\sigma_{12} - \bar{X}\sigma^2_1 - 2\bar{X}^3\sigma^2_2\\
& \sigma(b'_0,b'_2) = \sigma_{02} - \bar{X}\sigma_{12} + \bar{X}^2\sigma^2_2 \\
& \sigma(b'_1,b'_2) = \sigma_{12} - 2\bar{X}\sigma^2_2
\end{aligned}
$$

where $\sigma^2_0 = \sigma^2b_0, \sigma_{01} = \sigma b_0,b_1$, etc

# 8.34 
In a regression study, three types of banks were involved, namely, commercial, mutual savings, and savings and loan. Consider the following system of indicator variables for type of bank:

|Type of Bank| $X_2$| $X_3$|
|----------|------|------|
|Commercial |1| 0| 
|Mutual Savings |0| 1|
|Savings and loan |-1| -1|

## a. 
Develop a first-order linear regression model for relating last year's profit or loss ($Y$) to size of bank ($X_1$) and type of bank ($X_2, X_3$).

$Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + \beta_3X_{i3} + \epsilon$

## b. State the response functions for the three types of banks.

Commercial ($X_2=1, X_3=0$): 

$$
\begin{aligned}
E[Y] &= \beta_0 + \beta_1X_{1} + \beta_2(1) + \beta_3(0) \\
&= (\beta_0 + \beta_2) + \beta_1X_1
\end{aligned}
$$

Mutual savings ($X_2=0, X_3=1$): 

$$
\begin{aligned}
E[Y] &= \beta_0 + \beta_1X_{1} + \beta_2(0) + \beta_3(1) \\
&=  (\beta_0 + \beta_3) + \beta_1X_1
\end{aligned}
$$

Savings and loan ($X_2=-1, X_3=-1$): 

$$
\begin{aligned}
E[Y] &= \beta_0 + \beta_1X_{1} + \beta_2(-1) + \beta_3(-1) \\
&=  (\beta_0 - \beta_2 - \beta_3) + \beta_1X_1
\end{aligned}
$$

## c.
Interpret each of the following quantities: (1) $\beta_2$, (2) $\beta_3$, (3) $-\beta_2 - \beta_3$

(1) $\beta_2$ shows how much higher profit the commercial banks yield from the response function than for other banks of any size firm

(2) $\beta_3$ shows how much higher profit the mutual savings banks yield from the response function than for other banks of any size firm

(3) $- \beta_2 - \beta_3$ shows how much lower profit the savings loan banks yield from the response function than for other two banks of any size firm

# 8.39
Refer to the CDI data set in Appendix C.2. The number of active physicians ($Y$) is to be regressed against total population ($X_1$), total personal income ($X_2$), and geographic region ($X_3, X_4, X_5$).

## a. 
Fit a first-order regression model. Let $X_3 = 1$ if NE and 0 otherwise, $X_4 = 1$ if NC and 0 otherwise, and $X_5 = 1$ if South and 0 otherwise. In the event they are all 0, then it is West.

```{r}
CDI <- read.csv("CDI_Data.csv", header = F)

names(CDI) <- c("ID", "county", "state", "land_area", "total_pop", "precent_pop_18_34", "percent_pop_65", "num_physicians", "n_hospital_beds", "total_crimes", "percent_hs_grads", "percent_bach", "percent_pov", "percent_unemploy", "income_per_capita", "total_income", "geographic_region")

CDI$X3 <- ifelse(CDI$geographic_region == 1,1,0)
CDI$X4 <- ifelse(CDI$geographic_region == 2,1,0)
CDI$X5 <- ifelse(CDI$geographic_region == 3,1,0)
```

```{r}
mod <- lm(num_physicians ~ total_pop + total_income + X3 + X4 + X5, data = CDI)
summary(mod)
```

First order regression model:
$\hat{Y} = -207.5 + 0.0005X_1 + 0.107X_2 + 149X_3 + 145.5X_4 + 191.2X_5$

## b.
Examine whether the effect for the northeastern region on number of active physicians differs from the effect for the northeastern region by constructing an appropriate 90 percent confidence interval. Interpret your interval estimate.

General formula:

$b_k \pm t(\frac{1- \infty}{2}; n-p)s(b_k)$

For comparison:

$$
\begin{aligned}
& (b_3-b_4) \pm t(\frac{1- \infty}{2}; n-p)s(b_3-b_4) \\
& 149 - 145.5 \pm t(\frac{1- 0.1}{2}; 440-6)*(86.83-85.15) \\
& 3.5 \pm t(0.95;434) * 1.68 \\
& 3.5 \pm 1.645 * 1.68 \\
\end{aligned}
$$
$$
\begin{aligned}
& 3.5 + 1.645(1.68) = 6.26 \approx 6.3 \\
& 3.5 - 1.645(1.68) = 0.73
\end{aligned}
$$

## c. 
Test whether any geographic effects are present;'use $\alpha = .10$. State the alternatives, decision rule, and conclusion. What is the P-value of the test?

Alternatives:

$H_0:\beta_3 = \beta_4 = \beta_5 = 0$
$H_a: not \space all \space \beta_k = 0, (where \space k = 3,4,5)$

General Decision Rule
$$
\begin{aligned}
& F^* \leq F(1-\alpha, df_R-df_F, df_F),\;\text{fail to reject}\; H_0 \\
& F^* > F(1-\alpha, df_R-df_F, df_F), \;\text{reject}\; H_0
\end{aligned}
$$

```{r}
anova(mod)
nrow(CDI)
summary(mod)
```

$$
\begin{aligned}
& SSR(X_3,X_4,X_5|X_1,X_2) = 21097 + 23046 + 1829483 = 1873626 \\
& SSE(X_1,X_2,X_3,X_4,X_5) = 139093455
\end{aligned}
$$

$$
\begin{aligned}
F^* &= \frac{\frac{SSR(X_3,X_4,X_5|X_1,X_2)}{3}}{\frac{SSE(X_1,X_2,X_3,X_4,X_5)}{n-6}} \\
& = \frac{\frac{1873626}{3}}{\frac{139093455}{440-6}} \\
&= 1.9487
\end{aligned}
$$

```{r}
# 1-a, df, n-p (440-6)
qf(1-0.1, 3, 434)
```

$$
1.9487 \le 2.096449, \;\text{fail to reject}\; H_0
$$

```{r}
# another way to test full and reduced model
f_mod <- lm(num_physicians ~ total_pop + total_income + X3 + X4 + X5, data = CDI)
r_mod <- lm(num_physicians ~ total_pop + total_income, data = CDI)
anova(r_mod,f_mod, test="F") 
```

P-value of F* test = 0.121

























